Modularizing Street Map Generation \\ \vspace{8pt} \large An Adventure in Tensors

### 2013-12-12 paper/paper.tex
fix grammar in abstract
made citation style a bit more readable

\documentclass[twocolumn]{article}
\usepackage{amsmath}

@@

\begin{abstract}
This paper addresses the issue of creating modular, extensible code given an
existing solution to a problem, namely, interactively modeling large street
networks. We describe a functional implementation of a framework that allows
users to create a realistic street network over an empty domain, based on the
techniques described
in [Chen]. In particular, we were able to express the underlying mathematics
more intuitively than is possible in an imperative language.

@@

\section{Introduction}
Visual artists often face the time-consuming task of generating extensive,
realistic scenes by hand for movies and video games. However, the apparent
randomness found in systems like street layouts can be modelled effectively
with a variety of mathematical algorithms, giving rise to the growing field of
procedural generation. In this paper we will discuss an algorithm using tensor
fields, from [Chen], to generate realistic city street maps. To facilitate the
generation of aesthetically pleasing maps which fit an artist's requirements,
we implement a simple system of user-specified constraints to allow the user to
shape the resulting street map to fit their needs.

@@

\section{Mathematical Background}
A tensor $t$ is a *magical thing* represented by a 2x2 symmetric, traceless
matrix:
\[
    \sqmat{\cos{2\theta}}{\sin{2\theta}}{\sin{2\theta}}{-\cos{2\theta}}
    = \sqmat{a}{b}{b}{-a}
\]
where $R$ is the magnitude and $\theta$ is the direction, and
$R\geq0 \textrm{ and } \theta\in[0,2\pi)$. Since the matrix is symmetric and
traceless, it is only necessary to store two values, $a$ and $b$ to encode
the tensor.

@@

\subsection{Tensor Field Generation}
From each constraint, a tensor field is defined across the entire domain; this
is called a basis field. The basis fields are then combined to produce a tensor
field that describes the space and takes into account all of the specified
constraints. This is done using an exponential decay combinator, such that when
a basis is combined with others, the constraint that defined it has less
influence on the direction of the field the further a point is from it. The
equations in this section follow closely the system detailed in \cite{chen}.

@@

The first is self-explanatory; if the streamline leaves the boundary of the
map, then we don’t need to draw it any more. The second is one of the
alterations we used, based on the work of \cite{chen}. When designing a set of
roads, it is rare that two streets will approach each other and not connect,
whether they approach at a larger angle (an intersection) or are running mostly
parallel to each other (a merge zone). To approximate this, at each point we
find the nearest neighbor out of the streamlines that have already been drawn.
If it is within a certain separation distance, we draw a line directly between
the two points.

@@

\section{Assessment}
There were two main goals that we set out to achieve with this project. Here we
present a mostly heuristic, slightly quantitative attempt to qualify our
success, using both internal evaluation and comparisons to existing systems
\cite{chen}.

@@

\bibliographystyle{plain}
\bibliography{references}

>>>>

\documentclass[twocolumn]{article}
\usepackage{amsmath}
% \usepackage{natbib}

@@

\begin{abstract}
This paper addresses the issue of creating modular, extensible code given an
existing solution to a problem, namely, interactively modeling large street
networks. We describe a functional implementation of a framework that allows
users to create a realistic street network over an empty domain, based on the
techniques described
in [Chen]. In particular, we were able to express the underlying mathematics
more intuitively and clearly than is possible in an imperative language.

@@

\section{Introduction}
Visual artists often face the time-consuming task of generating extensive,
realistic scenes by hand for movies and video games. However, the apparent
randomness found in systems like street layouts can be modelled effectively
with a variety of mathematical algorithms, giving rise to the growing field of
procedural generation. In this paper we will discuss an algorithm using tensor
fields, from \cite{Chen}, to generate realistic city street maps. To facilitate
the generation of aesthetically pleasing maps which fit an artist's
requirements, we implement a simple system of user-specified constraints to
allow the user to shape the resulting street map to fit their needs.

@@

\section{Mathematical Background}
A tensor $t$ is a geometric object represented by a 2x2 symmetric, traceless
matrix:
\[
    \sqmat{\cos{2\theta}}{\sin{2\theta}}{\sin{2\theta}}{-\cos{2\theta}}
    = \sqmat{a}{b}{b}{-a}
\]
where $R$ is the magnitude and $\theta$ is the direction, and
$R\geq0 \textrm{ and } \theta\in[0,2\pi)$. Since the matrix is symmetric and
traceless, it is only necessary to store two values, $a$ and $b$ to encode
the tensor.

@@

\subsection{Tensor Field Generation}
From each constraint, a tensor field is defined across the entire domain; this
is called a basis field. The basis fields are then combined to produce a tensor
field that describes the space and takes into account all of the specified
constraints. This is done using an exponential decay combinator, such that when
a basis is combined with others, the constraint that defined it has less
influence on the direction of the field the further a point is from it. The
equations in this section follow closely the system detailed in \cite{Chen}.

@@

The first is self-explanatory; if the streamline leaves the boundary of the
map, then we don’t need to draw it any more. The second is one of the
alterations we used, based on the work of \cite{Chen}. When designing a set of
roads, it is rare that two streets will approach each other and not connect,
whether they approach at a larger angle (an intersection) or are running mostly
parallel to each other (a merge zone). To approximate this, at each point we
find the nearest neighbor out of the streamlines that have already been drawn.
If it is within a certain separation distance, we draw a line directly between
the two points.

@@

\section{Assessment}
There were two main goals that we set out to achieve with this project. Here we
present a mostly heuristic, slightly quantitative attempt to qualify our
success, using both internal evaluation and comparisons to existing systems
\cite{Chen}.

@@

\bibliographystyle{alpha}
\bibliography{references}
